### 2021-02-23 sections/3-results.tex
setup

\section{Results}
\label{sec:results}

Write about results \dots