### 2021-02-23 sections/4-discussion.tex

\section{Discussion}
\label{sec:discussion}

Discuss results and future work \dots

Include figures~\ref{fig:discussion:umap-annthyroid-euclidean}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=2.5in]{images/umaps/annthyroid-euclidean-umap2d.png}
    \caption{UMAP embedding of Annthyroid data with the Euclidean distance function.}
    \label{fig:discussion:umap-annthyroid-euclidean}
\end{figure}