### 2021-03-09 sections/3-results.tex
wrote outline

\section{Results}
\label{sec:results}

Write about results \dots

\subsection{Asymptotic Complexity}
\label{subsec:results:asymptotic-complexity}

Analyze:
\begin{itemize}
    \item Building the tree.
    \item $\rho$-nearest neighbors search.
    \item k-nearest neighbors search.
    \item compression:
    \begin{itemize}
        \item encode (novel contribution)
        \item serialize
        \item compress
        \item decompress
        \item desrialize
        \item decode (novel contribution)
    \end{itemize}
\end{itemize}

\subsection{Benchmarks}
\label{subsec:results:benchmarks}

Datasets and Distance function pairs, measure:

Linear search vs accelerated search runtime performance, speedup factor, number of distance comparisons.
Only at max-depth. leave out benchmarks at iteratively increasing depths.

Compression CAKES + gzip vs just gzip. Only at max-depth.