### 2021-03-09 sections/4-discussion.tex

\section{Discussion}
\label{sec:discussion}

Discuss results and future work \dots

CLAM-CAKES.
This is most effective when datasets exhibit low metric entropy and low local fractal dimension.

Unlike LSH, this is extensible to any user provided distance function.

Exactness of search and metric vs non-metric.

\subsection{Future Works}
\label{subsec:results:future-works}

GPU-acceleration for the more computationally intensive distance functions like wasserstein.

Live-updates as new points are added to the dataset.
For example, build tree with a subsample of the full data and then add the rest of the instances to simulate live-updates.

More distance functions. For example, maximal-common-subgraph for molecular structures.

TODO: Provide link to source code on github.