### 2021-04-07 sections/5-results.tex

\section{Results}
\label{sec:results}

Write about results \dots

\subsection{Asymptotic Complexity}
\label{subsec:results:asymptotic-complexity}

Analyze:
\begin{itemize}
    \item Building the tree.
    \item $\rho$-nearest neighbors search.
    \item k-nearest neighbors search.
    \item compression:
    \begin{itemize}
        \item encode (novel contribution)
        \item serialize
        \item compress
        \item decompress
        \item deserialize
        \item decode (novel contribution)
    \end{itemize}
\end{itemize}

\subsection{Benchmarks}
\label{subsec:results:benchmarks}

Specs of Ark, and maybe M1 MacBook Air and Dell XPS.

Provide link to source code on GitHub.

For each pair $(X, f)$ and a small selection of depths, measure:
\begin{enumerate}[1.]
    \item linear search vs accelerated search:
    \begin{enumerate}[i.]
        \item runtime performance,
        \item speedup factor, and
        \item number of distance comparisons.
    \end{enumerate}
    \item Compression CAKES + zlib vs just zlib, also substitute other libraries for zlib:
    \begin{enumerate}[i.]
        \item just the encoding vs the full dataset,
        \item encoding + serialization, and
        \item encoding + serialization + zlib.
    \end{enumerate}
\end{enumerate}

Find the optimal clusters for compression.
Compare linear search vs accelerated search terminated at these optimal clusters.