### 2021-04-07 sections/3-methods.tex

\section{Methods}
\label{sec:methods}

Write about methods and provide asymptotic complexity analysis of each method.

\subsection{Partition}
\label{subsec:methods:partition}

Partition algorithm for building the tree.
Lift this from the CHAODA paper.

Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.

\subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
\label{subsec:methods:rnn-search}

Modified binary search.
Lift this from CHESS paper.

Complexity is the same as in CHESS.

\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
\label{subsec:methods:knn-search}

Repeated $\rho$-NN search with where radius is modified with each iteration as guided by lfd of query ball.

Lift from CHESS paper as a starting point.

Complexity is some factor times CHESS-search.

\subsection{Compression}
\label{subsec:methods:compression}

Encode and Decode algorithms.
These could be, and probably will be, different for each $(X, f)$ pairing.

Asymptotic complexity analysis for:
\begin{enumerate}[i.]
    \item encode (novel contribution)
    \item serialize
    \item compress
    \item decompress
    \item deserialize
    \item decode (novel contribution)
\end{enumerate}

% discrete vs continuous, and small alphabet vs large alphabet in discrete