### 2021-04-07 sections/5-results.tex

\section{Results}
\label{sec:results}

Write about results \dots

Specs of Ark, and maybe M1 MacBook Air and Dell XPS.

Provide link to source code on GitHub.

For each pair $(X, f)$ and a small selection of depths, measure:
\begin{enumerate}[1.]
    \item linear search vs accelerated search:
    \begin{enumerate}[i.]
        \item runtime performance,
        \item speedup factor, and
        \item number of distance comparisons.
    \end{enumerate}
    \item Compression CAKES + zlib vs just zlib, also substitute other libraries for zlib:
    \begin{enumerate}[i.]
        \item just the encoding vs the full dataset,
        \item encoding + serialization, and
        \item encoding + serialization + zlib.
    \end{enumerate}
    \item Measure time taken for accelerated-search in the compressed space and decompression of the hits with different combinations of serde and codec libraries.
\end{enumerate}

Find the optimal clusters for compression.
Compare linear search vs accelerated search terminated at these optimal clusters.