### 2023-08-18 sections/3-methods.tex
struggles with intro and methods

\section{Methods}
\label{sec:methods}

Write about methods and provide asymptotic complexity analysis of each method.

\subsection{Dataset and Distance Function}
\label{subsec:methods:dataset-and-distance-function}

We start with a dataset $\textbf{X} = \{x_1 \dots x_n\}$ with $n$ points and a distance function $f : (\textbf{X}, \textbf{X}) \mapsto \mathbb{R}^+$.
Given two points from the dataset, the distance function deterministically returns a non-negative real number. We require that the distance function be symmetric ($f(x, y) = f(y, x) \forall x, y \in \textbf{X}$) and 
that the distance between two points $x$ and $y$ be zero if and only if $x = y$. When, in addition to these constraints, we also require that the distance function obey 
the triangle inequality (i.e., when the distance function is a metric), we can guarantee that our search algorithms have perfect recall.

CLAM assumes the ``manifold hypothesis''~\cite{fefferman2016testing}, which states that datasets collected from constrained generating phenomena that are embedded in a high-dimensional space typically only occupy a low-dimensional manifold in that space.
In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$. While we often Euclidean notions, such as voids and volumes, to talk about geometric and topological 
properties of clusters and of the manifold, CLAM does not rely on such notions; they serve merely as a convenient, more intuitive way to talk about the underlying mathematics.

\subsection{Clustering}
\label{subsec:methods:clustering}

We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with better selection of 
poles. For a cluster with $m$ points, we begin by computing a 
random sample of $\sqrt m$ points. We then compute the geometric median of this sample, which we call the 
cluster's center. We then compute the distance from the center to each point in the sample of $\sqrt m$ points. 
The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
from $l$ is designated the right pole. We then partition the cluster into a left child and a right child, where the 
left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
points in the cluster which are closer to $r$ than to $l$. Without loss of generality, we assign to the left child 
those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
repeat this procedure until each leaf contains only one datum or until some other user-specified stopping criterion 
is met.

\begin{algorithm} % enter the algorithm environment
\caption{Partition} % give the algorithm a caption
\label{alg:partition} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $cluster$
    \STATE $m \leftarrow \lfloor \sqrt{|cluster.points|} \rfloor$
    \STATE $seeds \leftarrow m$ random points from $cluster.points$
    \STATE $c \leftarrow$ geometric median of $seeds$
    \STATE $l \leftarrow \argmax d(c,x) \ \forall \ x \in cluster.points$
    \STATE $r \leftarrow \argmax d(l,x) \ \forall \ x \in cluster.points$
    \STATE $left \leftarrow \{x | x \in cluster.points \land d(l,x) \le d(r,x)\}$
    \STATE $right \leftarrow \{x | x \in cluster.points \land d(r,x) < d(l,x)\}$
    \IF{$|left| > 1$}
        \STATE Partition($left$)
    \ENDIF
    \IF{$|right| > 1$}
        \STATE Partition($right$)
    \ENDIF
\end{algorithmic}
\end{algorithm}

\subsubsection {Proof of Eventually Decreasing Cluster Radii}
\subsection{Dataset and Distance Function}
\label{subsec:methods:dataset-and-distance-function}

We start with a dataset $\textbf{X} = \{x_1 \dots x_n\}$ with $n$ points and a distance function $f : (\textbf{X}, \textbf{X}) \mapsto \mathbb{R}^+$.
Given two points from the dataset, the distance function deterministically returns a non-negative real number. We require that the distance function be symmetric ($f(x, y) = f(y, x) \forall x, y \in \textbf{X}$) and 
that the distance between two points $x$ and $y$ be zero if and only if $x = y$. When, in addition to these constraints, we also require that the distance function obey 
the triangle inequality (i.e., when the distance function is a metric), we can guarantee that our search algorithms have perfect recall.

CLAM assumes the ``manifold hypothesis''~\cite{fefferman2016testing}, which states that datasets collected from constrained generating phenomena that are embedded in a high-dimensional space typically only occupy a low-dimensional manifold in that space.
In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$. While we often Euclidean notions, such as voids and volumes, to talk about geometric and topological 
properties of clusters and of the manifold, CLAM does not rely on such notions; they serve merely as a convenient, more intuitive way to talk about the underlying mathematics.

\subsection{Clustering}
\label{subsec:methods:clustering}

We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with better selection of 
poles. For a cluster with $m$ points, we begin by computing a 
random sample of $\sqrt m$ points. We then compute the geometric median of this sample, which we call the 
cluster's center. We then compute the distance from the center to each point in the sample of $\sqrt m$ points. 
The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
from $l$ is designated the right pole. We then partition the cluster into a left child and a right child, where the 
left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
points in the cluster which are closer to $r$ than to $l$. Without loss of generality, we assign to the left child 
those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
repeat this procedure until each leaf contains only one datum or until some other user-specified stopping criterion 
is met.

\begin{algorithm} % enter the algorithm environment
\caption{Partition} % give the algorithm a caption
\label{alg:partition} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $cluster$
    \STATE $m \leftarrow \lfloor \sqrt{|cluster.points|} \rfloor$
    \STATE $seeds \leftarrow m$ random points from $cluster.points$
    \STATE $c \leftarrow$ geometric median of $seeds$
    \STATE $l \leftarrow \argmax d(c,x) \ \forall \ x \in cluster.points$
    \STATE $r \leftarrow \argmax d(l,x) \ \forall \ x \in cluster.points$
    \STATE $left \leftarrow \{x | x \in cluster.points \land d(l,x) \le d(r,x)\}$
    \STATE $right \leftarrow \{x | x \in cluster.points \land d(r,x) < d(l,x)\}$
    \IF{$|left| > 1$}
        \STATE Partition($left$)
    \ENDIF
    \IF{$|right| > 1$}
        \STATE Partition($right$)
    \ENDIF
\end{algorithmic}
\end{algorithm}

\subsubsection {Proof of Eventually Decreasing Cluster Radii}
\label{subsubsec:methods:clustering:proof-of-eventually-decreasing-cluster-radii}

\subsubsection {Complexity}
\label{subsubsec:methods:clustering:complexity}

Clustering: Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.

\subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
\label{subsec:methods:rnn-search}

Given a query $q$ and a search radius $\rho$, find all $x \in X$ s.t. $f(q, x) \leq \rho$.

Modified binary search to identify candidate leaf clusters.
Exhaustive leaf search over those leaves.

Lift this from CHESS paper. Be sure to cite CHESS here.

Complexity is the same as in CHESS.

\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
\label{subsec:methods:knn-search}

The k-nearest neighbors search problem is posed as follows: given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.

Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $k$-nn variants make use of the following definitions:
\begin{itemize}
    \item $\delta = f(q, c)$ is the distance from the query to the cluster center.
    \item $\delta_{max} = \delta + r$ is the distance from the query to the potentially farthest instance in the Cluster.
    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the potentially closest instance in the Cluster.
\end{itemize}

\subsubsection{Expanding Threshold}
\label{subsubsec:methods:knn-search:expanding-threshold}
In this variant of KNN, we repeatedly

Let $Q$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $Q$ with the root cluster.
Let $P$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.

While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
than $\delta_{min}$ of the top priority element in $Q$, do the following:
\begin{enumerate}
\item While the top priority element $q$ in $Q$ isn't a leaf, replace $q$ with its children.
\item Remove $q$ from $Q$ and add all of its points to $P$. 
\item Remove points from $P$ until $P$ has $k$ points. 
\end{enumerate}

Return $P$.

\begin{algorithm} % enter the algorithm environment
\caption{Expanding Threshold} % give the algorithm a caption
\label{alg:expanding_threshold} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $tree$, $query$, $k$
    \STATE $Q \leftarrow$ priority queue
    \STATE $Q.push(tree.root)$
    \STATE $P \leftarrow$ priority queue
    \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
        \WHILE{$!Q.min.isLeaf$}
            \STATE $[l, r] \leftarrow Q.extractMin().children()$
            \STATE $Q.push([l, r])$
        \ENDWHILE
        \STATE $leaf \leftarrow Q.extractMin()$
        \STATE $P.push(leaf)$
        \WHILE{$|P| > k$}
            \STATE $P.extractMax()$
        \ENDWHILE
    \ENDWHILE
    \STATE Return $P$
\end{algorithmic}
\end{algorithm}

\subsubsection{Sieve V1}
\label{subsubsec:methods:knn-search:sieve-v1}

In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.

Let $Q$ be a vector of clusters sorted by non-decreasing $\delta_{max}$. Initialize $Q$ with  
the root cluster. 
Let $P$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.

Repeat the following: 
\begin{enumerate}
\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
\item Remove from $P$ any points which are a distance greater than $\tau$ from the query. 
\item If no cluster in $Q$ has a $\delta_{max}$ less than or equal to $\tau$, break.
\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
\item Transfer from $Q$ to $P$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
\item Replace all non-leaf clusters in $Q$ with their children. 
\end{enumerate}

At this stage, if any clusters remain in $Q$, we add all their points to $P$. 
Extract $P$.

\begin{algorithm} % enter the algorithm environment
    \caption{Sieve V1} % give the algorithm a caption
    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[2] % enter the algorithmic environment
        \REQUIRE $tree$, $query$, $k$
        \STATE $Q \leftarrow$ priority queue
        \STATE $Q.push(tree.root)$
        \STATE $P \leftarrow$ priority queue
        \WHILE{$|P| < k \lor P.max.\delta > Q.max.\delta_{min}$}
            \WHILE{$!Q.extractMax().isLeaf$}
                \STATE $[l, r] \leftarrow Q.extractMax().children()$
                \STATE $Q.push([l, r])$
            \ENDWHILE
            \STATE $leaf \leftarrow Q.extractMax()$
            \STATE $P.push(leaf)$
            \WHILE{$|P| > k$}
                \STATE $P.extractMax()$
            \ENDWHILE
        \ENDWHILE
        \STATE Return $P$
    \end{algorithmic}
    \end{algorithm}

\subsubsection{Sieve V2}
\label{subsubsec:methods:knn-search:sieve-v2}
Add explanation of the sieve v2 algorithm.

\subsubsection{Repeated Rnn}
\label{subsubsec:methods:knn-search:repeated-rnn}
Add explanation of the repeated rnn algorithm.

\subsubsection{Leaf Search}
\label{subsubsec:methods:knn-search:leaf-search}

Let $hits$ be a a fixed size priority queue with $k$ instances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
Initialize $hits$ with the centers of $candidates$.

Let $f_k$ be the distance to the $k^{th}$ farthest instance so far.
Exclude all $candidates$ for which $\delta^2 > f_k$, i.e. their closest possible instance is farthest that the $k^{th}$ farthest instance found so far.
Remove the closest cluster from $candidates$ and add all its contained instances to $hits$.
Repeat until $candidates$ is empty.

\subsubsection{Complexity}
\label{subsubsec:methods:knn-search:complexity}

This is at-least on-par with the complexity for $\rho$-nearest neighbors search from CHESS.
It is potentially better because we effectively shrink the search radius to adapt to the $k$ closest instances found during the algorithm.

\label{subsubsec:methods:clustering:proof-of-eventually-decreasing-cluster-radii}

\subsubsection {Complexity}

\subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
\label{subsec:methods:rnn-search}

Given a query $q$ and a search radius $\rho$, find all $x \in X$ s.t. $f(q, x) \leq \rho$.

Modified binary search to identify candidate leaf clusters.
Exhaustive leaf search over those leaves.

Lift this from CHESS paper. Be sure to cite CHESS here.

Complexity is the same as in CHESS.

\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
\label{subsec:methods:knn-search}

The k-nearest neighbors search problem is posed as follows: given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.

Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $k$-nn variants make use of the following definitions:
\begin{itemize}
    \item $\delta = f(q, c)$ is the distance from the query to the cluster center.
    \item $\delta_{max} = \delta + r$ is the distance from the query to the potentially farthest instance in the Cluster.
    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the potentially closest instance in the Cluster.
\end{itemize}

\subsubsection{Expanding Threshold}
\label{subsubsec:methods:knn-search:expanding-threshold}
In this variant of KNN, we repeatedly

Let $Q$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $Q$ with the root cluster.
Let $P$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.

While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
than $\delta_{min}$ of the top priority element in $Q$, do the following:
\begin{enumerate}
\item While the top priority element $q$ in $Q$ isn't a leaf, replace $q$ with its children.
\item Remove $q$ from $Q$ and add all of its points to $P$. 
\item Remove points from $P$ until $P$ has $k$ points. 
\end{enumerate}

Return $P$.

\begin{algorithm} % enter the algorithm environment
\caption{Expanding Threshold} % give the algorithm a caption
\label{alg:expanding_threshold} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $tree$, $query$, $k$
    \STATE $Q \leftarrow$ priority queue
    \STATE $Q.push(tree.root)$
    \STATE $P \leftarrow$ priority queue
    \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
        \WHILE{$!Q.min.isLeaf$}
            \STATE $[l, r] \leftarrow Q.extractMin().children()$
            \STATE $Q.push([l, r])$
        \ENDWHILE
        \STATE $leaf \leftarrow Q.extractMin()$
        \STATE $P.push(leaf)$
        \WHILE{$|P| > k$}
            \STATE $P.extractMax()$
        \ENDWHILE
    \ENDWHILE
    \STATE Return $P$
\end{algorithmic}
\end{algorithm}

\subsubsection{Sieve V1}
\label{subsubsec:methods:knn-search:sieve-v1}

In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.

Let $Q$ be a vector of clusters sorted by non-decreasing $\delta_{max}$. Initialize $Q$ with  
the root cluster. 
Let $P$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.

Repeat the following: 
\begin{enumerate}
\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
\item Remove from $P$ any points which are a distance greater than $\tau$ from the query. 
\item If no cluster in $Q$ has a $\delta_{max}$ less than or equal to $\tau$, break.
\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
\item Transfer from $Q$ to $P$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
\item Replace all non-leaf clusters in $Q$ with their children. 
\end{enumerate}

At this stage, if any clusters remain in $Q$, we add all their points to $P$. 
Extract $P$.

\begin{algorithm} % enter the algorithm environment
    \caption{Sieve V1} % give the algorithm a caption
    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[2] % enter the algorithmic environment
        \REQUIRE $tree$, $query$, $k$
        \STATE $Q \leftarrow$ priority queue
        \STATE $Q.push(tree.root)$
        \STATE $P \leftarrow$ priority queue
        \WHILE{$|P| < k \lor P.max.\delta > Q.max.\delta_{min}$}
            \WHILE{$!Q.extractMax().isLeaf$}
                \STATE $[l, r] \leftarrow Q.extractMax().children()$
                \STATE $Q.push([l, r])$
            \ENDWHILE
            \STATE $leaf \leftarrow Q.extractMax()$
            \STATE $P.push(leaf)$
            \WHILE{$|P| > k$}
                \STATE $P.extractMax()$
            \ENDWHILE
        \ENDWHILE
        \STATE Return $P$
    \end{algorithmic}
    \end{algorithm}

\subsubsection{Sieve V2}
\label{subsubsec:methods:knn-search:sieve-v2}
Add explanation of the sieve v2 algorithm.

\subsubsection{Repeated Rnn}
\label{subsubsec:methods:knn-search:repeated-rnn}
Add explanation of the repeated rnn algorithm.

\subsubsection{Leaf Search}
\label{subsubsec:methods:knn-search:leaf-search}

Let $hits$ be a a fixed size priority queue with $k$ instances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
Initialize $hits$ with the centers of $candidates$.

Let $f_k$ be the distance to the $k^{th}$ farthest instance so far.
Exclude all $candidates$ for which $\delta^2 > f_k$, i.e. their closest possible instance is farthest that the $k^{th}$ farthest instance found so far.
Remove the closest cluster from $candidates$ and add all its contained instances to $hits$.
Repeat until $candidates$ is empty.

\subsubsection{Complexity}
\label{subsubsec:methods:knn-search:complexity}

Clustering: Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.

This is at-least on-par with the complexity for $\rho$-nearest neighbors search from CHESS.
It is potentially better because we effectively shrink the search radius to adapt to the $k$ closest instances found during the algorithm.