### 2025-06-24 sections/4-results.tex
reorganizing and moving to supplement

\section{Results}
\label{sec:results}

% The algorithms presented in this paper rely heavily on the local fractal dimension (LFD) being much smaller than than the embedding dimension of the dataset.
% As such, we begin by examining our assumptions about the LFD of real-world datasets (see Section~\ref{sec:results:lfd-of-datasets} and Figure~\ref{fig:results:lfd-plots}).

We compare the performance of the CAKES algorithms against HNSW, ANNOY and FAISS-IVF on the Fashion-MNIST, Glove-25, Sift and Random datasets.
We also compare how this performance scales to the augmented versions of these datasets.
We also report the performance of CAKES on the Silva and RadioML datasets (see Section~\ref{sec:results:scaling-behavior-and-recall}, Figure~\ref{fig:results:scaling-plots}, and Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random}).

% Finally, we test our intuitions about unbalanced clustering being better for search
% than balanced clustering (see Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations} and Figure~\ref{fig:results:distance-counts}).

% \subsection{Local Fractal Dimension of Datasets}
% \label{sec:results:lfd-of-datasets}

% Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks.
% Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-MNIST, Glove-25, Sift, Random, Silva 18S, and Radio-ML.
% In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD.
% This is because our algorithms scale exponentially in the LFD, a

% The Fashion-MNIST dataset has an embedding dimension of 784 and uses the Euclidean distance metric.
% In Figure~\ref{fig:results:fashion-mnist-lfd} we observe that until approximately depth 5, Fashion-MNIST's LFD is low (i.e., less than 4).
% It then starts increasing, reaching a peak of about 6 near depth 20, before decreasing to 1 at the maximum depth.

% The Glove-25 dataset has an embedding dimension of 25 and uses the cosine distance function which, notably, is not a metric.
% Relative to Fashion-MNIST, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
% All percentile lines for Glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
% In particular, Glove-25's LFD is less than 3 for all depths.

% The Sift dataset has an embedding dimension of 128 and uses the Euclidean distance metric.
% Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-MNIST and Glove-25.
% It increases sharply to a peak of 9 around a depth of 10.
% It then decreases smoothly until reaching the deepest leaves in the tree.

% We generated the Random dataset to have the same cardinality and dimensionality as Sift.
% We used a uniform distribution in a 128-dimensional unit-hypercube to generate the points and the Euclidean metric to measure distances among them.
% Figure~\ref{fig:results:random-lfd} shows that the character of this dataset is significantly different from the others.
% The LFD starts at 20 at depth 0 and all percentile lines decrease linearly with depth until reaching the leaves of the tree.
% The spread in LFD starts very small for the first few clusters and increases as depth increases.
% The LFD of approximately 20 for the root cluster $\mathcal{R}$ is what we expect for this random dataset.
% To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, i.e.,\,the fact that in high dimensional spaces, the minimum and maximum pairwise distances between any two points are approximately equal.
% As a result, $\mathcal{R}$'s radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
% A consequence of this is that, with high probability, for every point in $\mathcal{R}$, its distance from $c$ is greater than $\tfrac{r}{2}$;
% in other words, $B(c, \tfrac{r}{2})$ contains only $c$ while $B(c, r)$ contains the entire dataset.
% Given our definition of LFD in Equation~\ref{eq:methods:lfd-half}, this means that the LFD of $\mathcal{R}$ is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
% Theoretically, the LFD of this dataset should be 128, i.e.\, it should be the same as the embedding dimension.
% This reflects the difference between how we empirically measure the LFD and the value we would expect.
% With sample sizes larger than 1,000,000, we would expect the LFD to approach 128 until we have sampled $2^{128}$ points, at which point the LFD would be 128 and would stay at 128 for even larger sample sizes.
% Unfortunately, such a large sample is practically impossible to generate.

% The Silva-18S dataset consists of genomic sequences whose unaligned lengths are at-most 3,712.
% As such the embedding dimension is 3,712, though as discussed previously, the embedding dimension would be 50,000 in a multiple sequence alignment.
% We use the Levenshtein edit distance (a metric) to measure distances between sequences.
% This dataset, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
% In particular, LFD is less than 3 for all depths, hovering near 1 for clusters at depth 40 and deeper.

% The Radio-ML dataset consists of measurements of radio-frequency signals using 1,024 dimensional complex-valued vectors.
% We use the Dynamic Time Warping distance metric on this dataset.
% This dataset is synthetic~\cite{oshea2018radioml} but uses a far more elaborate generation process than our Random dataset.
% The LFD values show three distinct peaks around an LFD of 12 at or near depths of 8, 25 and 50.
% Each peak is followed by a linear decrease until encountering a sharp spike for the next peak.
% Within each of the tree portions, this dataset has a character very similar to that of the Random dataset.
% This suggests that the dataset obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth.
% This is likely the result of a piecewise uniform sampling strategy used to generate the different modulation modes present in the dataset.

% \begin{figure}
%     \captionsetup[subfigure]{aboveskip=-15pt,belowskip=-3pt}
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.99\textwidth]{images/lfd/fashion-mnist.png}\\
%         \subcaption{Fashion-MNIST}
%         \label{fig:results:fashion-mnist-lfd}
%     \end{subfigure}%
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.99\textwidth]{images/lfd/glove-25.png}\\
%         \subcaption{Glove-25}
%         \label{fig:results:glove-25-lfd}
%     \end{subfigure}
%     \\
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.99\textwidth]{images/lfd/sift.png}\\
%         \subcaption{Sift}
%         \label{fig:results:sift-lfd}
%     \end{subfigure}%
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.99\textwidth]{images/lfd/random.png}\\
%         \subcaption{A random dataset}
%         \label{fig:results:random-lfd}
%     \end{subfigure}
%     \\
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.99\textwidth]{images/lfd/silva-SSU-Ref.png}\\
%         \subcaption{Silva 18S}
%         \label{fig:results:silva-lfd}
%     \end{subfigure}%
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.99\textwidth]{images/lfd/radio-ml.png}\\
%         \subcaption{RadioML}
%         \label{fig:results:radioml-lfd}
%     \end{subfigure}%
%     \\
%     \vskip -0.1in
%     \begin{subfigure}[b]{0.94\textwidth}
%         \centering
%         \includegraphics[width=0.7\textwidth]{images/lfd/legend.png}
%         \label{fig:results:lfd-legend}
%     \end{subfigure}%
%     \vskip -0.1in
%     \caption{Local fractal dimension vs. cluster depth across six datasets. The `random' dataset is randomly generated according to the procedure in Section~\ref{sec:datasets-and-benchmarks:random-datasets}; note that the y-axis is different for this dataset. In each plot, the horizontal axis denotes depth in the cluster tree, and the vertical axis denotes the LFD of clusters at that depth. We show lines for the 5$^{th}$, 25$^{th}$, 50th, 75$^{th}$ and 95$^{th}$ percentiles of LFD, as well as the minimum and maximum LFD at each depth. So that plots best reflect the distribution of LFDs across the entire \textit{dataset}, we count each cluster as many times as its cardinality. For example, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 3.}
%     \label{fig:results:lfd-plots}
%     \vskip -0.4in
% \end{figure}

\subsection{Indexing and Tuning Time}
\label{sec:results:indexing-and-tuning-time}

For each of the ANN-benchmark datasets and the Random dataset, we report the time taken for each algorithm to build the index and to tune the hyper-parameters for these indices to achieve the highest possible recall. For the sake of brevity, these results are reported in the supplement.

\subsection{Scaling Behavior and Recall}
\label{sec:results:scaling-behavior-and-recall}

We benchmark the CAKES algorithms, na\"{i}ve linear search, HNSW, ANNOY, and FAISS-IVF on the Fashion-MNIST, Glove-25, Sift, and Random datasets.
We augment these datasets with synthetic points to examine how performance scales with cardinality.
We also benchmark the CAKES algorithms on the Silva and RadioML datasets, and we subsample these datasets instead of augmenting them to examine how performance scales with cardinality.
For each dataset, we use the corresponding distance function (see Table~\ref{tab:datasets:summary}).
For HNSW, ANNOY, and FAISS-IVF, we allow for a hyper-parameter search to tune their index for maximum recall.
For CAKES, we build the tree and use our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) to select the fastest algorithm for each dataset and cardinality.
We show results for $k=10$ in the main text, but results for $k=100$ can be found in the Supplement.
Figure~\ref{fig:results:scaling-plots} shows the scaling behavior of these algorithms on these datasets while Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random} report the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-MNIST, Glove-25, Sift, and Random respectively.

On the Fashion-MNIST, Glove-25, and Sift datasets (in Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} respectively), we observe that as cardinality increases, the Depth-First Sieve algorithm is consistently the fastest CAKES algorithm with a throughput that is constant in the cardinality of the dataset.
The Breadth-First Sieve algorithm is the usually the second fastest, also with a nearly constant throughput across all cardinalities.
The Repeated $\rho$-NN algorithm, however, falls off in throughput as cardinality increases.
All three CAKES algorithms exhibit perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset (which uses cosine distance).
In contrast, HNSW and ANNOY are faster than CAKES's algorithms for all cardinalities and have near constant throughput as cardinality increases, but their recall degrades quickly as cardinality increases.
FAISS-IVF exhibits linearly decreasing throughput as cardinality increases, which is expected from the algorithm given that we tune the hyper-parameters to maximize recall.

On the Random dataset, HNSW and ANNOY are still the fastest algorithms but exhibit recall values near 0.
CAKES's algorithms show linearly decreasing throughput as cardinality increases and are also slower than na\"{i}ve linear search.

Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random} show the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-MNIST, Glove-25, Sift, and Random respectively.
Though the plots in Figure~\ref{fig:results:scaling-plots} present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
We used our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) for each new tree (i.e.,\,at each cardinality), and this approach was always able to select the fastest algorithm for each dataset at each cardinality.
Since we also allow for tuning hyper-parameters for the other algorithms, and we allow for different sets of hyper-parameters at each cardinality, it is a fair comparison for these tables to only list the performance of the fastest CAKES algorithm.
When reporting recall, we use $1.000*$ to denote that the recall is imperfect, but rounds to $1.000$.

For Silva and RadioML, we benchmarked only CAKES's algorithms because HNSW, ANNOY and FAISS support neither the required distance functions nor, in the case of RadioML, complex-valued data.
Due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset---rather than augmented versions of the dataset---to examine how performance scales with cardinality.
For the Silva dataset (see Figure~\ref{fig:results:silva-scaling}), we observe that throughput initially decreases linearly as cardinality increases, but levels off at higher cardinalities.
Depth-First Sieve is, once again, the fastest CAKES algorithm.
For Radio-ML (see Figure~\ref{fig:results:radioml-scaling}), however, we see that throughput declines nearly linearly with cardinality, and that the three CAKES algorithms are, again, slower that na\"{i}ve linear search.

% \begin{figure}
%     \captionsetup[subfigure]{aboveskip=-5pt,belowskip=-0.5pt}
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics{plots/fashion-mnist_PermutedBall_10_throughput.png}
%         \subcaption{Fashion-MNIST for $k=10$.}
%         \label{fig:results:fashion-mnist-scaling}
%     \end{subfigure}%
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.95\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
%         \subcaption{Glove-25 for $k=10$.}
%         \label{fig:results:glove-25-scaling}
%     \end{subfigure}%
%     \\
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
%         \subcaption{Sift for $k=10$.}
%         \label{fig:results:sift-scaling}
%     \end{subfigure}%
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.95\textwidth]{plots/random_PermutedBall_10_throughput.png}
%         \subcaption{A random dataset for $k=10$.}
%         \label{fig:results:random-scaling}
%     \end{subfigure}%
%     \\
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.95\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
%         \subcaption{Silva for $k=10$.}
%         \label{fig:results:silva-scaling}
%     \end{subfigure}%
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.95\textwidth]{plots/radio-ml_Ball_10_throughput.png}
%         \subcaption{RadioML for $k=10$ at SnR = 10dB.}
%         \label{fig:results:radioml-scaling}
%     \end{subfigure}%
%     \\
%     \begin{subfigure}[b]{0.94\textwidth}
%         \centering
%         \includegraphics[width=0.7\textwidth]{plots/legend.png}
%         \label{fig:results:scaling-legend}
%     \end{subfigure}%
%     \caption{Throughput across six datasets, including a randomly-generated dataset.
%     In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
%     For Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms become faster than linear search but the cardinality at which this occurs differs by dataset.
%     For Fashion-MNIST and Glove-25, Depth-First Sieve is consistently fastest, while for Sift, Repeated $\rho$-NN is the fastest for smaller cardinalities and Depth-First Sieve is the fastest for larger cardinalities.
%     With Silva, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it starts to level off at higher cardinalities.
%     Depth-First Sieve is consistently the fastest algorithm on the Silva dataset.
%     For Radio-ML and Random, we see that all three CAKES algorithms are slower than na\"{i}ve linear search, and that their throughput decreases linearly with cardinality.
%     HNSW and ANNOY are the fastest algorithms on all four datasets we benchmarked them on, but their recall degrades quickly as cardinality increases on all datasets.
%     In particular, on the Random dataset, HNSW and ANNOY have near-zero recall.}
%     \label{fig:results:scaling-plots}
% \end{figure}

% \begin{table}
%     \caption{Throughput (QPS) and Recall on the Fashion-MNIST dataset.
%     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
%     In contrast, CAKES has perfect recall on Fashion-MNIST, since it used with Euclidean distance, which is a metric.}
%     \label{tab:results:qps-and-recall-fmn}
%     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
%         \hline
%         \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
%         & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
%         \cline{1-9}
%         \hline
%         1   & \num{1.33e4} & 0.954  & \num{2.19e3} & 0.950  & \num{2.01e3} & 1.000* & \num{3.46e3} & 1.000  \\\cline{1-9}
%         2   & \num{1.38e4} & 0.803  & \num{2.12e3} & 0.927  & \num{9.39e2} & 1.000* & \num{3.68e3} & 1.000  \\\cline{1-9}
%         4   & \num{1.66e4} & 0.681  & \num{2.04e3} & 0.898  & \num{4.61e2} & 0.997  & \num{3.44e3} & 1.000  \\\cline{1-9}
%         8   & \num{1.68e4} & 0.525  & \num{1.93e3} & 0.857  & \num{2.26e2} & 0.995  & \num{3.30e3} & 1.000  \\\cline{1-9}
%         16  & \num{1.87e4} & 0.494  & \num{1.84e3} & 0.862  & \num{1.17e2} & 0.991  & \num{3.34e3} & 1.000  \\\cline{1-9}
%         32  & \num{1.56e4} & 0.542  & \num{1.85e3} & 0.775  & \num{5.91e1} & 0.985  & \num{2.96e3} & 1.000  \\\cline{1-9}
%         64  & \num{1.50e4} & 0.378  & \num{1.78e3} & 0.677  & \num{2.61e1} & 0.968  & \num{3.25e3} & 1.000  \\\cline{1-9}
%         128 & \num{1.49e4} & 0.357  & \num{1.66e3} & 0.538  & \num{1.33e1} & 0.964  & \num{2.96e3} & 1.000  \\\cline{1-9}
%         256 & --           & --     & \num{1.60e3} & 0.592  & \num{6.65e0} & 0.962  & \num{2.79e3} & 1.000  \\\cline{1-9}
%         512 & --           & --     & \num{1.83e3} & 0.581  & \num{3.56e0} & 0.949 & \num{2.84e3} & 1.000 \\
%         \hline
%     \end{tabular}
%     \vskip -0.2in
% \end{table}

% \begin{table}
%     \caption{Throughput (QPS) and Recall on the Glove-25 dataset.
%     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
%     In contrast, CAKES has near-perfect recall on Glove-25, since it used with cosine distance, which is not a metric.}
%     \label{tab:results:qps-and-recall-glove}
%     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
%         \hline
%         \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
%         & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
%         \cline{1-9}
%         \hline
%         1   & \num{2.28e4} & 0.801 & \num{2.83e3} & 0.835 & \num{2.38e3} & 1.000* & \num{1.54e3} & 1.000* \\\cline{1-9}
%         2   & \num{2.38e4} & 0.607 & \num{2.70e3} & 0.832 & \num{1.19e3} & 1.000* & \num{1.49e3} & 1.000* \\\cline{1-9}
%         4   & \num{2.50e4} & 0.443 & \num{2.61e3} & 0.839 & \num{6.19e2} & 1.000* & \num{1.28e3} & 1.000* \\\cline{1-9}
%         8   & \num{2.78e4} & 0.294 & \num{2.51e3} & 0.834 & \num{3.03e2} & 1.000* & \num{1.30e3} & 1.000* \\\cline{1-9}
%         16  & \num{3.11e4} & 0.213 & \num{2.23e3} & 0.885 & \num{1.51e2} & 1.000* & \num{1.14e3} & 1.000* \\\cline{1-9}
%         32  & \num{3.24e4} & 0.178 & \num{2.01e3} & 0.764 & \num{7.40e1} & 0.999  & \num{1.05e3} & 1.000* \\\cline{1-9}
%         64  & --           & --    & \num{1.99e3} & 0.631 & \num{3.77e1} & 0.997  & \num{1.07e3} & 1.000* \\\cline{1-9}
%         128 & --           & --    & --           & --    & \num{1.90e1} & 0.998  & \num{8.92e2} & 1.000* \\\cline{1-9}
%         256 & --           & --    & --           & --    & \num{9.47e0} & 0.998  & \num{8.91e2} & 1.000* \\
%         \hline
%     \end{tabular}
%     \vskip -0.2in
% \end{table}

% \begin{table}
%     \caption{Throughput (QPS) and Recall on the Sift dataset.
%     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
%     In contrast, CAKES has perfect recall on Sift, since it is used with Euclidean distance, which is a metric.}
%     \label{tab:results:qps-and-recall-sift}
%     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
%         \hline
%         \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
%         & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
%         \cline{1-9}
%         \hline
%         1   & \num{1.93e4} & 0.782 & \num{3.98e3} & 0.686 & \num{6.98e2} & 1.000* & \num{6.20e2} & 1.000 \\\cline{1-9}
%         2   & \num{2.03e4} & 0.552 & \num{3.80e3} & 0.614 & \num{3.30e2} & 1.000* & \num{2.95e2} & 1.000 \\\cline{1-9}
%         4   & \num{2.18e4} & 0.394 & \num{3.69e3} & 0.637 & \num{1.65e2} & 1.000* & \num{1.76e2} & 1.000 \\\cline{1-9}
%         8   & \num{2.48e4} & 0.298 & \num{3.58e3} & 0.710 & \num{7.72e1} & 1.000* & \num{1.27e2} & 1.000 \\\cline{1-9}
%         16  & \num{2.68e4} & 0.210 & \num{3.50e3} & 0.690 & \num{3.98e1} & 1.000* & \num{1.47e2} & 1.000 \\\cline{1-9}
%         32  & \num{2.75e4} & 0.193 & \num{3.44e3} & 0.639 & \num{2.09e1} & 0.999  & \num{1.24e2} & 1.000 \\\cline{1-9}
%         64  & --           & --    & \num{3.39e3} & 0.678 & \num{8.87e0} & 0.997  & \num{1.34e2} & 1.000 \\\cline{1-9}
%         128 & --           & --    & \num{3.36e3} & 0.643 & \num{4.78e0} & 0.993  & \num{1.31e2} & 1.000 \\
%         \hline
%     \end{tabular}
%     \vskip -0.2in
% \end{table}

% \begin{table}
%     \caption{Throughput (QPS) and Recall on the Random dataset.
%     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, we observe that CAKES's algorithms perform quite slowly.
%     As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
%     HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets.
%     }
%     \label{tab:results:qps-and-recall-random}
%     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
%         \hline
%         \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
%         & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
%         \cline{1-9}
%         \hline
%         1  & \num{1.17e4} & 0.060 & \num{4.28e3} & 0.028 & \num{7.342} & 1.000* & \num{6.06e2} & 1.000 \\\cline{1-9}
%         2  & \num{1.01e4} & 0.048 & \num{4.04e3} & 0.021 & \num{3.582} & 1.000* & \num{2.75e2} & 1.000 \\\cline{1-9}
%         4  & \num{9.12e3} & 0.031 & \num{3.64e3} & 0.014 & \num{1.902} & 1.000* & \num{1.35e2} & 1.000 \\\cline{1-9}
%         8  & \num{8.35e3} & 0.022 & \num{3.37e3} & 0.013 & \num{8.841} & 1.000* & \num{6.13e1} & 1.000 \\\cline{1-9}
%         16 & \num{8.25e3} & 0.008 & \num{3.17e3} & 0.006 & \num{4.361} & 1.000* & \num{2.82e1} & 1.000 \\\cline{1-9}
%         32 & --           & --    & \num{3.01e3} & 0.007 & \num{1.721} & 1.000* & \num{1.31e1} & 1.000 \\
%         \hline
%     \end{tabular}
%     \vskip -0.2in
% \end{table}

% \begin{figure}
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
%         \subcaption{Repeated $\rho$-NN}
%         \label{fig:results:fashion-mnist-counts-throughput}
%     \end{subfigure}%
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}
%         \subcaption{Repeated $\rho$-NN}
%         \label{fig:results:glove-25-counts-counts}
%     \end{subfigure}%
%     \\
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
%         \subcaption{Breadth First Sieve}
%         \label{fig:results:sift-counts-throughput}
%     \end{subfigure}%
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
%         \subcaption{Breadth First Sieve}
%         \label{fig:results:random-counts-counts}
%     \end{subfigure}%
%     \\
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
%         \subcaption{Depth First Sieve}
%         \label{fig:results:silva-counts-throughput}
%     \end{subfigure}%
%     \begin{subfigure}[b]{0.5\textwidth}
%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
%         \subcaption{Depth First Sieve}
%         \label{fig:results:radioml-counts-counts}
%     \end{subfigure}%
%     \\
%     \begin{subfigure}[b]{0.94\textwidth}
%         \centering
%         \includegraphics[width=0.6\textwidth]{images/distance_counts/legend.png}
%         \label{fig:results:counts-legend}
%     \end{subfigure}%
%     \caption{Number of distance computations across four clustering strategies and three search algorithms on the Fashion-MNIST dataset.
%     Adding the instrumentation to count the number of distance computations had the side-effect of significantly slowing down the search algorithms compared to those reported in Figure~\ref{fig:results:scaling-plots}.
%     The left column shows the throughput in queries per second, while the right column shows the mean number of distance computations per query.
%     The x-axis represents increasing cardinality of the dataset.}
%     \label{fig:results:distance-counts}
% \end{figure}